### 2016-04-12 manuscript.tex
Rewrite algorithm section

\section{Algorithm} Directly computing $\hat{a}[d,d']$ over the set $d,d'\in
\{0,1,\dots,P\}^2$ would be cubic in the number of segregating sites, but as is
the case with ALDER, we can use using a fast Fourier transform (FFT) to
approximate $\hat{a}$, giving an algorithm whose run-time is instead linear in
the number of segregating sites. We first rearrange $\hat{a}$ to get
\begin{align*} \hat{a}[d,d'] &=\frac{n}{(n-1)(n-2)}
\frac{\sum_{i=1}^n\sum_{x,y,z\in S[d,d']}
\delta_x\delta_y\delta_z(H_{i,x}-f_x)(H_{i,y}-f_y)(H_{i,z}-f_z)} {\sum_{x,y,z\in
S[d,d']}1}, \end{align*} and define sequences $b_i[d]$ and $c[d]$ by binning the
data and then doubling the length by padding with $P$ zeros, \begin{align*}
b_i[d] &= \left\{ \begin{array}{ll} \sum_{x: d\leq \lfloor{x}\rfloor <
d+1}\delta_x(H_{i,x}-f_x) & : 0\leq d < P\\ 0 & : P\leq d <2P \end{array}
\right.\\ c[d] &= \left\{ \begin{array}{ll} |\{x: d\leq \lfloor{x}\rfloor <
d+1\}| & : 0\leq d < P\\ 0 & : P\leq d <2P \end{array} \right. \end{align*} We
can approximate $|S[d,d']|$ and the $n$ sums in the numerator of $\hat{a}[d,d']$
in terms of convolutions of these sequences: $$ |S[d,d']| \approx \sum_{w=0}^P
c[w]c[w+d]c[w+d+d'] $$ $$ \sum_{x,y,z\in
S[d,d']}\delta_x\delta_y\delta_z(H_{i,x}-f_x)(H_{i,y}-f_y)(H_{i,z}-f_z) \approx
\sum_{w=0}^P b_i[w]b_i[w+d]b_i[w+d+d']. $$ These convolutions can be efficiently
computed with an FFT, since under a two-dimensional discrete Fourier transform, $$
\sum_{w=0}^P b_i[w]b_i[w+d]b_i[w+d+d'] \leftrightarrow
B_i[j]\bar{B_i}[k]B_i[k-j], $$ where $B_i$ is the (one-dimensional) discrete
Fourier transform of $b$, and $B_i[-j]$ is the $j^\text{th}$ to last most
element of $B_i$. Summing over $i$ and taking the inverse discrete Fourier
transform, we can approximate the discrete Fourier transform of the numerator of
$\hat{a}$. We use the same method applied to $c$ to approximate the denominator
of $\hat{a}$. Because the number of bins is generally much less than the number
of segregating sites, the rate-limiting step of this algorithm is the binning
step to form $c$ and the $b_i$'s, which is $O(S_n)$, rather than the FFTs, which
are $O(P^2\log(P))$.

>>>>

\section{Algorithm}
Directly computing $\hat{a}[d,d']$ over the set $d,d'\in \{0,1,\dots,P\}^2$
would be cubic in the number of segregating sites. However, by using the fast
Fourier Transform (FFT) technique introduced in ALDER \cite{loh2013inferring},
we can approximate $\hat{a}$ with an algorithm whose time complexity is instead
linear in the number of segregating sites.

First, rearrange $\hat{a}$ to get
\begin{align*}
	\hat{a}[d,d'] &=
	\frac{n}{(n-1)(n-2)}
	\frac{
		\sum_{i=1}^n \sum_{x,y,z\in S[d,d']}
		\delta_x \delta_y \delta_z (H_{i,x} - f_x) (H_{i,y} - f_y) (H_{i,z} - f_z)
	}{
		\sum_{x,y,z \in S[d,d']}
		1
	},
\end{align*}
and define sequences $b_i[d]$ and $c[d]$ by binning the data and then doubling
the length by padding with $P$ zeros,
\begin{align*}
	b_i[d] &=
	\left\{
		\begin{array}{ll}
			\sum_{x: d\leq \lfloor{x}\rfloor < d+1} \delta_x(H_{i,x}-f_x) & : 0 \leq d < P\\
			0 & : P\leq d <2P
		\end{array}
	\right.\\
	c[d] &=
	\left\{
		\begin{array}{ll}
			|\{x: d\leq \lfloor{x}\rfloor < d+1\}| & : 0\leq d < P\\
			0 & : P\leq d <2P
		\end{array}
	\right.
\end{align*}
We can approximate $|S[d,d']|$ and the $n$ sums in the numerator of
$\hat{a}[d,d']$ in terms of convolutions of these sequences:
$$
	|S[d,d']| \approx \sum_{w=0}^P c[w]c[w+d]c[w+d+d']
$$
$$
	\sum_{x,y,z\in S[d,d']}
	\delta_x\delta_y\delta_z(H_{i,x}-f_x)(H_{i,y}-f_y)(H_{i,z}-f_z)
	\approx
	\sum_{w=0}^P b_i[w]b_i[w+d]b_i[w+d+d'].
$$
These convolutions can be efficiently
computed with an FFT, since under a two-dimensional discrete Fourier transform
from $(d, d')$-space to $(j, k)$-space,
$$
\sum_{w=0}^P b_i[w]b_i[w+d]b_i[w+d+d']
\leftrightarrow
B_i[j]\bar{B_i}[k]B_i[k-j],
$$
where $B_i$ is the one-dimensional discrete Fourier transform of $b$ and for
$j>0$, $B_i[-j]$ is the $j^\text{th}$ to last most element of $B_i$.
Summing over $i$ and taking the inverse discrete Fourier transform, we can
approximate the discrete Fourier transform of the numerator of $\hat{a}$. We
apply the same method to $c$ to approximate the denominator of $\hat{a}$.

The time complexities for the binning and the FFT's are $O(S_n)$ and
$O(P^2\log(P))$. Of these two, the first term will dominate, because the number
of bins much less than the number of segregating sites.
